A SAT-Based Counterexample Guided Method for Unbounded Synthesis

### 2016-01-26 paper.tex
commented out authors for double-blind reviewing
made a pass over background definitions

\documentclass{llncs}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{fixltx2e}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{tikz}
\usetikzlibrary{trees}
\usetikzlibrary{positioning}

@@

\author{Alexander Legg\inst{1} 
    \and Nina Narodytska\inst{2}
    \and Leonid Ryzhyk\inst{2}}

\institute{NICTA\thanks{NICTA is funded by the Australian Government as represented by the Department of Broadband,
    Communications and the Digital Economy and the Australian Research Council through the ICT
    Centre of Excellence program.} and UNSW \\
    \email{alexander.legg@nicta.com.au}
    \and Samsung Research America}

@@

@@

\section{Reactive Synthesis}

A \emph{safety game}, $G = \langle X, L_u, L_c, \delta, I, E, \rangle$,
consists of a set of boolean state variables, sets of uncontrollable and
controllable label variables, a transition relationship $\delta : (X, L_u, L_c)
\to X$, an initial state, and an error state. The \emph{controller} and
\emph{environment} players choose controllable and uncontrollable labels
respectively and the game proceeds according to $\delta$.

An \emph{run} of a game $(x_0, u_0, c_0), (x_1, u_1, c_1) \dots (x_n, u_n,
c_n)$ is a chain of state and label pairs of length $n$ s.t.  $x_{k+1}
\leftarrow \delta(x_k, u_k, c_k)$. A run is winning for the controller if $x_0
= I \land \forall i \in \{1..n\} (x_i \neq E)$. In a bounded game of rank $n$
all runs are restricted to length $n$, whereas unbounded games consider runs of
infinite length. Since we consider only deterministic games a run is equivalent
to a list of assignments to $L_c$ and $L_u$.

@@

An \emph{environment strategy} $\pi^e : X \to L_u$ is a mapping of states to
uncontrollable labels. A bounded run is winning for the environment if $x_0
= I \land \exists i \in \{1..n\} (x_i = E)$ and an environment strategy is
winning for a bounded game if there exists a run $(x_0, \pi^e(x_1), c_1), (x_1,
\pi^e(x_1), c_1) \dots (x_n, \pi^e(x_n), c_n)$ that wins for the environment.
Safety games are zero sum, therefore the existence of a controller strategy
implies the nonexistence of an environment strategy and vice versa.

\subsection{Abstract Game Trees}

@@

A strategy is equivalent to the set of all runs with the player's labels
obeying the strategy mapping $\pi$. Therefore, a strategy can also be
represented by a tree. Relaxing the restriction of the strategy mapping allows
for a \emph{partial strategy} in which multiple labels are now possible for a
single state.

@@

\subsection{Counterexample Guided Bounded Synthesis}

@@

This technique was proposed by Narodytska et al \cite{narodytska2014}. In
Section \ref{sect:unbounded} we propose an extension to synthesis of unbounded
games but the original algorithm is outlined here for completeness.

@@

One aspect of the algorithm proposed by Narodytska et al\cite{narodytska2014}
is that it mimics real-world uses of synthesis by allowing cooperation between
system and environment. Unfortunately, allowing the opponents to pick moves for
one another can frequently backfire. It is common when modelling real-world
systems to abstract over out of scope failures with transitions that
immediately determine the outcome of the game. For example, in a network driver
a request to send a packet may fail due to a disconnected wire, which might be
modelled as an environment controlled transition to a system-winning state.

>>>>

\documentclass{llncs}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{fixltx2e}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{trees}
\usetikzlibrary{positioning}

@@

\author{}
\institute{}

%\author{Alexander Legg\inst{1} 
%    \and Nina Narodytska\inst{2}
%    \and Leonid Ryzhyk\inst{2}}
%
%\institute{NICTA\thanks{NICTA is funded by the Australian Government as represented by the Department of Broadband,
%    Communications and the Digital Economy and the Australian Research Council through the ICT
%    Centre of Excellence program.} and UNSW \\
%    \email{alexander.legg@nicta.com.au}
%    \and Samsung Research America}

@@

\sloppy

@@

\section{Background}

A \emph{safety game}, $G = \langle X, L_u, L_c, \delta, I, E \rangle$,
is defined over boolean state variables $X$, uncontrollable label variables $L_u$, and
controllable label variables $L_c$.  $I$ is the initial state of the game given as a 
valuation of state variables.  $E(X)$ is the set of error states represented by its 
characteristic formula.  The transition relation $\delta(X, L_u, L_c, X')$ of the game 
is a boolean formula that relates current state and label to the set of possible next 
states of the game.  We assume deterministic games, where 
$\delta(x,u,c,x'_1) \land \delta(x,u,c,x'_2) \implies (x'_1=x'_2)$.

At every round of the game, the \emph{environment} picks an
uncontrollable label, the \emph{controller} responds by choosing a controllable 
label and the game transitions into a new state according to $\delta$. 
A \emph{run} of a game $(x_0, u_0, c_0), (x_1, u_1, c_1) \dots (x_n, u_n,
c_n)$ is a chain of state and label pairs s.t.\,$\delta(x_k, u_k, c_k, x_{k+1})$. 
A run is winning for the controller if $x_0 = I \land \forall i \in \{1..n\} (\lnot E(x_i))$. 
In a \emph{bounded game} of rank $n$ all runs are restricted to length $n$, whereas unbounded 
games consider runs of infinite length. Since we consider only deterministic games, a run 
is uniquely described by a list of assignments to $L_c$ and $L_u$.

@@

An \emph{environment strategy} $\pi^e : X \to L_u$ is a mapping of states to
uncontrollable labels. A bounded run is winning for the environment if $x_0
= I \land \exists i \in \{1..n\} (E(x_i))$ and an environment strategy is
winning for a bounded game if all runs $(x_0, \pi^e(x_1), c_1), (x_1,
\pi^e(x_1), c_1) \dots (x_n, \pi^e(x_n), c_n)$ are winning for the environment.
Safety games are zero sum, therefore the existence of a winning controller strategy
implies the nonexistence of a winning environment strategy and vice versa.

\subsection{Counterexample-guided bounded synthesis}

We review the bounded synthesis algorithm by Narodytska et al.~\cite{narodytska2014},
which is the main building block for our unbounded algorithm.

@@

A strategy is equivalent to the set of all runs with the controller's labels
obeying the strategy mapping $\pi$. Therefore, a strategy can also be
represented by a tree. Relaxing the restriction of the strategy mapping allows
for a \emph{partial strategy} in which multiple labels are now possible for a
single state.

@@

@@

@@

One aspect of the algorithm proposed by Narodytska et al.~\cite{narodytska2014}
is that it mimics real-world uses of synthesis by allowing cooperation between
system and environment. Unfortunately, allowing the opponents to pick moves for
one another can frequently backfire. It is common when modelling real-world
systems to abstract over out of scope failures with transitions that
immediately determine the outcome of the game. For example, in a network driver
a request to send a packet may fail due to a disconnected wire, which might be
modelled as an environment controlled transition to a system-winning state.
